---
title: "2017 PGA Tour Statistical Analysis"
author: "Ryan Kyaw"
date: "May 2, 2020"
output: html_document
---
Is there statistical significance that any Strokes Gained statistics directly correlates to lower golf scores?

```{r, echo = FALSE}
library(tidyverse)
library(ggpubr)
```

```{r, echo = FALSE}
df_2017 <- read.csv("2017.csv")
```

Looking at the data from the 2017 golf season, we can use a graphical analysis to see how strokes gained compares to average score. 

```{r}
ggplot(df_2017) +
  geom_point(aes(x = df_2017$SG_PUTTING_PER_ROUND, y = df_2017$AVG_SCORE), color = "blue") +
  ggtitle("SG:Putting vs Average Score")+
  xlab("SG:Putting")+
  ylab("Average Score")+
  geom_smooth(method = "lm", aes(x = df_2017$SG_PUTTING_PER_ROUND, y = df_2017$AVG_SCORE), color = "red")
```

```{r}
ggplot(df_2017) +
  geom_point(aes(x = df_2017$SG.OTT, y = df_2017$AVG_SCORE), color = "brown")+
  ggtitle("SG:OTT vs Average Score")+
  xlab("SG:OTT")+
  ylab("Average Score")+
  geom_smooth(method = "lm", aes(x = df_2017$SG.OTT, y = df_2017$AVG_SCORE), color = "blue")
```

```{r}
ggplot(df_2017) +
  geom_point(aes(x = df_2017$SG.APR, y = df_2017$AVG_SCORE), color = "red")+
  ggtitle("SG:APR vs Average Score")+
  xlab("SG:APR")+
  ylab("Average Score")+
  geom_smooth(method = "lm", aes(x = df_2017$SG.APR, y = df_2017$AVG_SCORE), color = "blue")
```

```{r}
ggplot(df_2017) +
  geom_point(aes(x = df_2017$SG.ARG, y = df_2017$AVG_SCORE), color = "purple")+
  ggtitle("SG:ARG vs Average Score")+
  xlab("SG:ARG")+
  ylab("Average Score")+
  geom_smooth(method = "lm", aes(x = df_2017$SG.ARG, y = df_2017$AVG_SCORE), color = "red")
```

```{r}
cor(x = df_2017$AVG_SCORE, y = df_2017$SG.APR)
cor(x = df_2017$AVG_SCORE, y = df_2017$SG_PUTTING_PER_ROUND)
cor(x = df_2017$AVG_SCORE, y = df_2017$SG.ARG)
cor(x = df_2017$AVG_SCORE, y = df_2017$SG.OTT)
```

The negative correlation values appear to imply that the higher SG values correlate to lower scores. 

Consider the density plots for each of our parameters.

```{r}
ggplot()+
  geom_density(aes(x = df_2017$AVG_SCORE), fill = "red", alpha = 0.3)
```

```{r}
ggplot()+
  geom_density(aes(x = df_2017$SG_PUTTING_PER_ROUND), fill = "blue", alpha = 0.3)+
  geom_density(aes(x = df_2017$SG.APR), fill = "brown", alpha = 0.3)+
  geom_density(aes(x = df_2017$SG.OTT), fill = "red", alpha = 0.3)+
  geom_density(aes(x = df_2017$SG.ARG), fill = "gold", alpha = 0.3)
```

Each of our parameters appear to be normally distributed. 

Let's create linear models for each of the Strokes Gained categories versus Scoring Average. 

```{r}
lm_putt <- lm(AVG_SCORE ~ SG_PUTTING_PER_ROUND, data = df_2017)
summary(lm_putt)
BIC(lm_putt)
```

```{r}
lm_drive <- lm(AVG_SCORE ~ SG.OTT, data = df_2017)
summary(lm_drive)
BIC(lm_drive)
```

```{r}
lm_ARG <- lm(AVG_SCORE ~ SG.ARG, data = df_2017)
summary(lm_ARG)
```

```{r}
lm_APR <- lm(AVG_SCORE ~ SG.APR, data = df_2017)
summary(lm_APR)
```

Let's run a hypothesis test to see if the correlation between the Strokes Gained variables and Average Score is statistically significant. 

Pearson Correlation Coefficient

```{r}
putttest <- cor.test(df_2017$SG_PUTTING_PER_ROUND, df_2017$AVG_SCORE, method = "pearson")
putttest
```

```{r}
argtest <- cor.test(df_2017$SG.ARG, df_2017$AVG_SCORE, method = "pearson")
argtest
```

```{r}
aprtest <- cor.test(df_2017$SG.APR, df_2017$AVG_SCORE, method = "pearson")
aprtest
```

```{r}
otttest <- cor.test(df_2017$SG.OTT, df_2017$AVG_SCORE, method = "pearson")
otttest
```

